 Allocation scoring rules as a generalization of quantile scoring rules

### 2023-05-01 alloscore_manuscript/allocation.tex
small comment

This scoring rule evaluates the forecast distribution $F$ only through its $\alpha$ quantile. While this is faithful to the context of the decision-making problem, it may not be satisfying as a measure of the quality of the full forecast distribution. For this purpose, one option is to integrate the quantile loss across different values of the quantile level $\alpha$, or, equivalently, across different values of the decision problem's parameters $C$ and $L$. This gives rise to the (weighted) CRPS:
\begin{align*}
    S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) p(C, L) \, dC dL
\end{align*}
The usual CRPS results from taking $p$ to be the density of a distribution such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.
The weighted interval score (WIS) can be used when the full forecast distribution $F$ is not available, and amounts to an approximation of CRPS where the distribution $p$ is discrete with point masses at the quantile levels corresponding to the endpoints of a finite set of prediction intervals.

>>>>

This scoring rule evaluates the forecast distribution $F$ only through its $\alpha$ quantile. While this is faithful to the context of the decision-making problem, it may not be satisfying as a measure of the quality of the full forecast distribution. For this purpose, one option is to integrate the quantile loss across different values of the quantile level $\alpha$, or, equivalently, across different values of the decision problem's parameters $C$ and $L$. This gives rise to the (weighted) CRPS:
\begin{align*}
    S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) p(C, L) \, dC dL
\end{align*}
\apg{Quick note that I don't think this will work without some more identifications. The integral needs to be over probability levels and thresholds and without the some discussion like the one commented out (currently) at line 123 or (better in my mind) one like I'm trying to do with binary choices at the unit-level, I don't think we have a threshold.}
The usual CRPS results from taking $p$ to be the density of a distribution such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.
The weighted interval score (WIS) can be used when the full forecast distribution $F$ is not available, and amounts to an approximation of CRPS where the distribution $p$ is discrete with point masses at the quantile levels corresponding to the endpoints of a finite set of prediction intervals.

### 2023-05-01 notes/allocation_notes.rmd
two very small but potentially crucial additions
small comment

Or if $F=\Phi_{\mu,\sigma}$ is normal with location $\mu = E_F[Y]$ and scale $\sigma = \sqrt{\mathrm{Var}(Y)}$, 
we have 
\begin{align}
\int_{-\infty}^{\frac{x-\mu}{\sigma}}zdF_0(z)=-\frac{1}{\sqrt{2\pi}}\exp\left\{-\frac{(x-\mu)^2}{2 \sigma^2}\right\} = - \sigma\varphi_{\mu, \sigma}(x)
\end{align}
so that
\begin{align}
\sbar_F(x) = \kappa\left[(\Phi_{\mu,\sigma}(x) - \alpha)(x-\mu) + \sigma^2 \varphi_{\mu, \sigma}(x)\right].
\end{align}

@@

Similarly, a weather forecaster might be judged by the cost $Cx$ of recommended protection $x$ against a level $y$ of adverse weather (e.g., rainfall) added to any loss $L(x-y)_-$ resulting from underprediction, leading to the scoring function 
\begin{align}
s_m(x,y) = Cx + L(x-y)_-,
\end{align}
which rearranges to 
\begin{align}
s_m(x,y) = C(x-y)_+ + (L-C)(x-y)_- + Cy.
\end{align}
The minimizer of $\sbar_{m,F}(x) = E_F[s_m(x,Y)]$ given the forecaster's belief $Y \sim F$ is now
\begin{align}
Q = F^{-1}\left(\frac{L-C}{C+L -C}\right) =F^{-1}\left(1-\frac{C}{L}\right).
\end{align}
In particular, faced with the classical binary decision problem of whether to recommend an additional unit of protection given a current level of protection $x$, the forecaster's optimal decision rule under $s_m$ is to recommend adding protection if $x < F^{-1}\left(1-\frac{C}{L}\right)$, that is, if 
\begin{align}
1-F(x) = \mathbb{P}_F\{y>x\} > \frac{C}{L},
\end{align}
the \emph{cost-loss ratio} of the problem.

>>>>

Or if $F=\Phi_{\mu,\sigma}$ is normal with location $\mu = E_F[Y]$ and scale $\sigma = \sqrt{\mathrm{Var}_F(Y)}$, 
we have 
\begin{align}
\int_{-\infty}^{\frac{x-\mu}{\sigma}}zdF_0(z)=-\frac{1}{\sqrt{2\pi}}\exp\left\{-\frac{(x-\mu)^2}{2 \sigma^2}\right\} = - \sigma\varphi_{\mu, \sigma}(x)
\end{align}
so that
\begin{align}
\sbar_F(x) = \kappa\left[(\Phi_{\mu,\sigma}(x) - \alpha)(x-\mu) + \sigma^2 \varphi_{\mu, \sigma}(x)\right].
\end{align}

@@

Similarly, a weather forecaster might be judged by the cost $Cx$ of recommended protection $x$ against a level $y$ of adverse weather (e.g., rainfall) added to any loss $L(x-y)_-$ resulting from underprediction, leading to the scoring function 
\begin{align}
s_m(x,y) = Cx + L(x-y)_-,
\end{align}
which rearranges to 
\begin{align}
s_m(x,y) = C(x-y)_+ + (L-C)(x-y)_- + Cy.
\end{align}
The minimizer of $\sbar_{m,F}(x) = E_F[s_m(x,Y)]$ given the forecaster's belief $Y \sim F$ is now
\begin{align}
Q = F^{-1}\left(\frac{L-C}{C+L -C}\right) =F^{-1}\left(1-\frac{C}{L}\right).
\end{align}

In particular, faced with the classical binary decision problem of whether to recommend an additional unit of protection given a current level of protection $x$, the forecaster's optimal decision rule under $s_m$ is to recommend adding protection if $x < F^{-1}\left(1-\frac{C}{L}\right)$, that is, if 
\begin{align}
1-F(x) = \mathbb{P}_F\{y>x\} > \frac{C}{L},
\end{align}
the \emph{cost-loss ratio} of the problem.
